---
title: "Ex. 8.1"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures-8-1/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T)
```

### Parameters

```{r par}
dir <- "~/git/variani/study/01-book-Kruschke-DBDA/doc/programs/DBDA2Eprograms"
```

## Include DBDA2 files

```{r inc, cache = F}
library(coda)
```

## Example `Jags-Ydich-Xnom1subj-MbernBeta-Example.R`

See code in the `Jags-Ydich-Xnom1subj-MbernBeta-Example.R` file.

### Source files

```{r src, echo = F, results = 'hide'}
dir0 <- getwd()
dir <- "~/git/variani/study/01-book-Kruschke-DBDA/doc/programs/DBDA2Eprograms"
f <- "Jags-Ydich-Xnom1subj-MbernBeta.R"

setwd(dir)
source(f)
setwd(dir0)

# for plotting in Rmd by diagMCMC function
openGraph <- function(...) return(invisible())
```

Sourced files: ``r paste(f, collapse = ", ")``.

### Parameters

```{r par1}
fn <- "Jags-Ydich-Xnom1subj-MbernBeta-" 

nsteps <- 10000
```

### Read data

```{r dat}
dat <- read.csv(file.path(dir, "z15N50.csv"))
str(dat)
```

```{r dat_tab}
table(dat$y)
```

### MCMC samples

```{r mcmc, results = 'hide', cache = T}
myData <- dat
out <- genMCMC(data = myData, numSavedSteps = nsteps, saveName = fn)
```

```{r mcmc_sum}
summary(out)
```

### Plot diagnostics

```{r diagMCMC, cache = T}
diagMCMC(out)
```

### Plot posteriors

```{r post, cache = T}
plotMCMC(out, data = myData)
```

## Example `Jags-Ydich-XnomSsubj-MbernBeta-Example.R`

See code in the `Jags-Ydich-XnomSsubj-MbernBeta-Example.R` file.

### Source files

```{r src2, echo = F, results = 'hide'}
dir0 <- getwd()
f <- "Jags-Ydich-XnomSsubj-MbernBeta.R"

setwd(dir)
source(f)
setwd(dir0)

# for plotting in Rmd by diagMCMC function
openGraph <- function(...) return(invisible())
```

Sourced files: ``r paste(f, collapse = ", ")``.

### Parameters

```{r par2}
fn <- "Jags-Ydich-XnomSsubj-MbernBeta-"

nsteps <- 50000
```

### Read data

```{r dat2}
dat <- read.csv(file.path(dir, "z6N8z2N7.csv"))
str(dat)
```

### MCMC samples

```{r mcmc2, results = 'hide', cache = T}
myData <- dat
out <- genMCMC(data = myData, numSavedSteps = nsteps, saveName = fn)
```

```{r mcmc_sum2}
summary(out)
```

### Plot diagnostics

```{r diagMCMC2, cache = T}
l_ply(varnames(out), function(p) {
  diagMCMC(out, p)
})
```

### Plot posteriors

```{r post2, cache = T}
plotMCMC(out, data = myData, compVal = NULL , compValDiff = 0.0, ropeDiff = c(-0.05, 0.05))
```

## Exercise 8.1: two subjets with different sample size

### Read data

```{r dat3}
dat <- read.csv(file.path(dir, "z6N8z2N7.csv"))

# remove some samples from `dat` for the Subject #2
dat <- rbind(dat[1:8, ], dat[1:8, ], dat[1:8, ], dat[1:8, ], dat[1:8, ],
  dat[9:15, ])

str(dat)
table(dat$s)
```

### MCMC samples

```{r mcmc3, results = 'hide', cache = T}
myData <- dat 
out <- genMCMC(data = myData, numSavedSteps = nsteps, saveName = fn)
```

```{r mcmc_sum3}
summary(out) 
```

### Plot diagnostics

```{r diagMCMC3, cache = T}
l_ply(varnames(out), function(p) {
  diagMCMC(out, p) 
})
```

### Plot posteriors

```{r post3, cache = T}
plotMCMC(out, data = myData, compVal = NULL , compValDiff = 0.0, ropeDiff = c(-0.05, 0.05)) 
```

### Conclusions

More samples for the Subject #1 improves estimation of the theta 1 parameter,
as well as estimation of the difference between thetas 1 and 2.

In principle, a better estimation of theta 2 is not expected, as the number of samples remained the same in the Example.
There is also no the shrinkage effect, because the model is not hierachical.

## Exercise 8.1: three subjets with different sample size

### Read data

Use the same data set given in the Solutions.

```{r dat4}
# remove some samples from `dat` for the Subject #2
dat <- data.frame(y = c( rep(1,9), rep(0,3), rep(1,45), rep(0,15), rep(1,3),rep(0,9)),
  s = c(rep("A",12), rep("B",60), rep("C",12)))

str(dat)
table(dat$s)
```

### MCMC samples

```{r mcmc4, results = 'hide', cache = T}
myData <- dat 
out <- genMCMC(data = myData, numSavedSteps = nsteps, saveName = fn)
```

```{r mcmc_sum4}
summary(out) 
```

### Plot diagnostics

```{r diagMCMC4, cache = T}
l_ply(varnames(out), function(p) {
  diagMCMC(out, p) 
})
```

### Plot posteriors

```{r post4, cache = T}
plotMCMC(out, data = myData, compVal = NULL , compValDiff = 0.0, ropeDiff = c(-0.05, 0.05)) 
```

### Conclusions

The large sample size (theta 2 vs. theta 1/3) the more narrow the HDI interval.

## Exercise 8.2

```{r smry, cache = T}
smryMCMC(out, rope = c(0.4, 0.6), ropeDiff = c(-0.1, 0.1))
```

### Conclusions

The summary output provides numbers to quantify what is plotted by the plot functions.

## Exercise 8.4

### genMCMC1

Changes in the code of `genMCMC`:

```
genMCMC1 <- function(...)
{
# ...
 dataList = list(
    #y = y ,
    s = s ,
    Ntotal = Ntotal ,
    Nsubj = Nsubj
  )
# ...  
}
```

```{r genMCMC1, echo = F}
genMCMC1 <- function( data , numSavedSteps=50000 , saveName=NULL ) { 
  require(rjags)
  #-----------------------------------------------------------------------------
  # THE DATA.
  # N.B.: This function expects the data to be a data frame, 
  # with one component named y being a vector of integer 0,1 values,
  # and one component named s being a factor of subject identifiers.
  y = data$y
  s = as.numeric(data$s) # converts character to consecutive integer levels
  # Do some checking that data make sense:
  if ( any( y!=0 & y!=1 ) ) { stop("All y values must be 0 or 1.") }
  Ntotal = length(y)
  Nsubj = length(unique(s))
  # Specify the data in a list, for later shipment to JAGS:
  dataList = list(
    #y = y ,
    s = s ,
    Ntotal = Ntotal ,
    Nsubj = Nsubj
  )
  #-----------------------------------------------------------------------------
  # THE MODEL.
  modelString = "
  model {
    for ( i in 1:Ntotal ) {
      y[i] ~ dbern( theta[s[i]] )
    }
    for ( sIdx in 1:Nsubj ) {
      theta[sIdx] ~ dbeta( 2 , 2 ) # N.B.: 2,2 prior; change as appropriate.
    }
  }
  " # close quote for modelString
  writeLines( modelString , con="TEMPmodel.txt" )
  #-----------------------------------------------------------------------------
  # INTIALIZE THE CHAINS.
  # Initial values of MCMC chains based on data:
  # Option 1: Use single initial value for all chains:
  #  thetaInit = rep(0,Nsubj)
  #  for ( sIdx in 1:Nsubj ) { # for each subject
  #    includeRows = ( s == sIdx ) # identify rows of this subject
  #    yThisSubj = y[includeRows]  # extract data of this subject
  #    thetaInit[sIdx] = sum(yThisSubj)/length(yThisSubj) # proportion
  #  }
  #  initsList = list( theta=thetaInit )
  # Option 2: Use function that generates random values near MLE:
  initsList = function() {
    thetaInit = rep(0,Nsubj)
    for ( sIdx in 1:Nsubj ) { # for each subject
      includeRows = ( s == sIdx ) # identify rows of this subject
      yThisSubj = y[includeRows]  # extract data of this subject
      resampledY = sample( yThisSubj , replace=TRUE ) # resample
      thetaInit[sIdx] = sum(resampledY)/length(resampledY) 
    }
    thetaInit = 0.001+0.998*thetaInit # keep away from 0,1
    return( list( theta=thetaInit ) )
  }
  #-----------------------------------------------------------------------------
  # RUN THE CHAINS
  parameters = c( "theta")     # The parameters to be monitored
  adaptSteps = 500             # Number of steps to adapt the samplers
  burnInSteps = 500            # Number of steps to burn-in the chains
  nChains = 4                  # nChains should be 2 or more for diagnostics 
  thinSteps = 1
  nIter = ceiling( ( numSavedSteps * thinSteps ) / nChains )
  # Create, initialize, and adapt the model:
  jagsModel = jags.model( "TEMPmodel.txt" , data=dataList , inits=initsList , 
                          n.chains=nChains , n.adapt=adaptSteps )
  # Burn-in:
  cat( "Burning in the MCMC chain...\n" )
  update( jagsModel , n.iter=burnInSteps )
  # The saved MCMC chain:
  cat( "Sampling final MCMC chain...\n" )
  codaSamples = coda.samples( jagsModel , variable.names=parameters , 
                              n.iter=nIter , thin=thinSteps )
  # resulting codaSamples object has these indices: 
  #   codaSamples[[ chainIdx ]][ stepIdx , paramIdx ]
  if ( !is.null(saveName) ) {
    save( codaSamples , file=paste(saveName,"Mcmc.Rdata",sep="") )
  }
  return( codaSamples )
}
```

```{r run1, results = 'hide', cache = T}
dat <- read.csv(file.path(dir, "z6N8z2N7.csv"))
myData <- dat
out <- genMCMC1(data = myData)
```

```{r plot_run1, cache = T}
plotMCMC(out, data = myData)
```

### genMCMC2

Changes in the code of `genMCMC`:

```
genMCMC2 <- function(...)
{
# ...
 dataList = list(
    #y = y ,
    s = s ,
    Ntotal = Ntotal ,
    Nsubj = Nsubj
  )
# ...  

  modelString = "
  model {
    for ( i in 1:Ntotal ) {
      y[i] ~ dbern( theta[s[i]] )
    }
    for ( sIdx in 1:Nsubj ) {
      theta[sIdx] ~ dbeta( 1 , 1 ) # N.B.: 2,2 prior; change as appropriate.
    }
  }
  "
}
```

```{r genMCMC2, echo = F}
genMCMC2 <- function( data , numSavedSteps=50000 , saveName=NULL ) { 
  require(rjags)
  #-----------------------------------------------------------------------------
  # THE DATA.
  # N.B.: This function expects the data to be a data frame, 
  # with one component named y being a vector of integer 0,1 values,
  # and one component named s being a factor of subject identifiers.
  y = data$y
  s = as.numeric(data$s) # converts character to consecutive integer levels
  # Do some checking that data make sense:
  if ( any( y!=0 & y!=1 ) ) { stop("All y values must be 0 or 1.") }
  Ntotal = length(y)
  Nsubj = length(unique(s))
  # Specify the data in a list, for later shipment to JAGS:
  dataList = list(
    #y = y ,
    s = s ,
    Ntotal = Ntotal ,
    Nsubj = Nsubj
  )
  #-----------------------------------------------------------------------------
  # THE MODEL.
  modelString = "
  model {
    for ( i in 1:Ntotal ) {
      y[i] ~ dbern( theta[s[i]] )
    }
    for ( sIdx in 1:Nsubj ) {
      theta[sIdx] ~ dbeta( 1 , 1 ) # N.B.: 2,2 prior; change as appropriate.
    }
  }
  " # close quote for modelString
  writeLines( modelString , con="TEMPmodel.txt" )
  #-----------------------------------------------------------------------------
  # INTIALIZE THE CHAINS.
  # Initial values of MCMC chains based on data:
  # Option 1: Use single initial value for all chains:
  #  thetaInit = rep(0,Nsubj)
  #  for ( sIdx in 1:Nsubj ) { # for each subject
  #    includeRows = ( s == sIdx ) # identify rows of this subject
  #    yThisSubj = y[includeRows]  # extract data of this subject
  #    thetaInit[sIdx] = sum(yThisSubj)/length(yThisSubj) # proportion
  #  }
  #  initsList = list( theta=thetaInit )
  # Option 2: Use function that generates random values near MLE:
  initsList = function() {
    thetaInit = rep(0,Nsubj)
    for ( sIdx in 1:Nsubj ) { # for each subject
      includeRows = ( s == sIdx ) # identify rows of this subject
      yThisSubj = y[includeRows]  # extract data of this subject
      resampledY = sample( yThisSubj , replace=TRUE ) # resample
      thetaInit[sIdx] = sum(resampledY)/length(resampledY) 
    }
    thetaInit = 0.001+0.998*thetaInit # keep away from 0,1
    return( list( theta=thetaInit ) )
  }
  #-----------------------------------------------------------------------------
  # RUN THE CHAINS
  parameters = c( "theta")     # The parameters to be monitored
  adaptSteps = 500             # Number of steps to adapt the samplers
  burnInSteps = 500            # Number of steps to burn-in the chains
  nChains = 4                  # nChains should be 2 or more for diagnostics 
  thinSteps = 1
  nIter = ceiling( ( numSavedSteps * thinSteps ) / nChains )
  # Create, initialize, and adapt the model:
  jagsModel = jags.model( "TEMPmodel.txt" , data=dataList , inits=initsList , 
                          n.chains=nChains , n.adapt=adaptSteps )
  # Burn-in:
  cat( "Burning in the MCMC chain...\n" )
  update( jagsModel , n.iter=burnInSteps )
  # The saved MCMC chain:
  cat( "Sampling final MCMC chain...\n" )
  codaSamples = coda.samples( jagsModel , variable.names=parameters , 
                              n.iter=nIter , thin=thinSteps )
  # resulting codaSamples object has these indices: 
  #   codaSamples[[ chainIdx ]][ stepIdx , paramIdx ]
  if ( !is.null(saveName) ) {
    save( codaSamples , file=paste(saveName,"Mcmc.Rdata",sep="") )
  }
  return( codaSamples )
}
```

```{r run2, results = 'hide', cache = T}
dat <- read.csv(file.path(dir, "z6N8z2N7.csv"))
myData <- dat
out <- genMCMC2(data = myData)
```

```{r plot_run2, cache = T}
plotMCMC(out, data = myData)
```

### genMCMC3

Changes in the code of `genMCMC`:

```
genMCMC3 <- function(...)
{
# ...
 dataList = list(
    #y = y ,
    s = s ,
    Ntotal = Ntotal ,
    Nsubj = Nsubj
  )
# ...  

  modelString = "
  model {
    for ( i in 1:Ntotal ) {
      y[i] ~ dbern( theta[s[i]] )
    }
    for ( sIdx in 1:Nsubj ) {
      theta[sIdx] ~ dbeta( 0.5 , 0.5 ) # N.B.: 2,2 prior; change as appropriate.
    }
  }
  "
}
```

```{r genMCMC3, echo = F}
genMCMC3 <- function( data , numSavedSteps=50000 , saveName=NULL ) { 
  require(rjags)
  #-----------------------------------------------------------------------------
  # THE DATA.
  # N.B.: This function expects the data to be a data frame, 
  # with one component named y being a vector of integer 0,1 values,
  # and one component named s being a factor of subject identifiers.
  y = data$y
  s = as.numeric(data$s) # converts character to consecutive integer levels
  # Do some checking that data make sense:
  if ( any( y!=0 & y!=1 ) ) { stop("All y values must be 0 or 1.") }
  Ntotal = length(y)
  Nsubj = length(unique(s))
  # Specify the data in a list, for later shipment to JAGS:
  dataList = list(
    #y = y ,
    s = s ,
    Ntotal = Ntotal ,
    Nsubj = Nsubj
  )
  #-----------------------------------------------------------------------------
  # THE MODEL.
  modelString = "
  model {
    for ( i in 1:Ntotal ) {
      y[i] ~ dbern( theta[s[i]] )
    }
    for ( sIdx in 1:Nsubj ) {
      theta[sIdx] ~ dbeta( 0.5 , 0.5 ) # N.B.: 2,2 prior; change as appropriate.
    }
  }
  " # close quote for modelString
  writeLines( modelString , con="TEMPmodel.txt" )
  #-----------------------------------------------------------------------------
  # INTIALIZE THE CHAINS.
  # Initial values of MCMC chains based on data:
  # Option 1: Use single initial value for all chains:
  #  thetaInit = rep(0,Nsubj)
  #  for ( sIdx in 1:Nsubj ) { # for each subject
  #    includeRows = ( s == sIdx ) # identify rows of this subject
  #    yThisSubj = y[includeRows]  # extract data of this subject
  #    thetaInit[sIdx] = sum(yThisSubj)/length(yThisSubj) # proportion
  #  }
  #  initsList = list( theta=thetaInit )
  # Option 2: Use function that generates random values near MLE:
  initsList = function() {
    thetaInit = rep(0,Nsubj)
    for ( sIdx in 1:Nsubj ) { # for each subject
      includeRows = ( s == sIdx ) # identify rows of this subject
      yThisSubj = y[includeRows]  # extract data of this subject
      resampledY = sample( yThisSubj , replace=TRUE ) # resample
      thetaInit[sIdx] = sum(resampledY)/length(resampledY) 
    }
    thetaInit = 0.001+0.998*thetaInit # keep away from 0,1
    return( list( theta=thetaInit ) )
  }
  #-----------------------------------------------------------------------------
  # RUN THE CHAINS
  parameters = c( "theta")     # The parameters to be monitored
  adaptSteps = 500             # Number of steps to adapt the samplers
  burnInSteps = 500            # Number of steps to burn-in the chains
  nChains = 4                  # nChains should be 2 or more for diagnostics 
  thinSteps = 1
  nIter = ceiling( ( numSavedSteps * thinSteps ) / nChains )
  # Create, initialize, and adapt the model:
  jagsModel = jags.model( "TEMPmodel.txt" , data=dataList , inits=initsList , 
                          n.chains=nChains , n.adapt=adaptSteps )
  # Burn-in:
  cat( "Burning in the MCMC chain...\n" )
  update( jagsModel , n.iter=burnInSteps )
  # The saved MCMC chain:
  cat( "Sampling final MCMC chain...\n" )
  codaSamples = coda.samples( jagsModel , variable.names=parameters , 
                              n.iter=nIter , thin=thinSteps )
  # resulting codaSamples object has these indices: 
  #   codaSamples[[ chainIdx ]][ stepIdx , paramIdx ]
  if ( !is.null(saveName) ) {
    save( codaSamples , file=paste(saveName,"Mcmc.Rdata",sep="") )
  }
  return( codaSamples )
}
```

```{r run3, results = 'hide', cache = T}
dat <- read.csv(file.path(dir, "z6N8z2N7.csv"))
myData <- dat
out <- genMCMC3(data = myData)
```

```{r plot_run3, cache = T}
plotMCMC(out, data = myData)
```

### Conlusion

Priors on individual parameters has little to deal with the prior on the difference of parameters.
