---
title: "Ex. 12"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures-12/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T)
```

## Include 

### Include `dbda`

```{r src, cache = F}
load_all("~/git/variani/dbda/")
```

## Exercise 12.1

```{r data}
z <- 7
N <- 24
```

### Part A

```{r 1_A}
theta <- 0.5

p <- theta^z * (1 - theta)^(N - z)
p
```

### Part B

```{r 1_B, cache = T}
# spike prior on 0.5
a <- 2000
b <- 2000

BernBeta(c(a, b), c(rep(0, N-z), rep(1, z)) , 
  ROPE = c(0.48, 0.52),
  plotType = "Bars", showCentTend = "Mode", showHDI = TRUE, showpD = TRUE)
```

The resulted p-value, as shown on the figure:

```{r}
pDnull <- 6.02e-8
```

### Part C
 
 
```{r 1_C, cache = T}
# Haldane prior
a <- 0.01
b <- 0.01

BernBeta(c(a, b), c(rep(0, N-z), rep(1, z)) , 
  ROPE = c(0.48, 0.52),
  plotType = "Bars", showCentTend = "Mode", showHDI = TRUE, showpD = TRUE)
```

The resulted p-value, as shown on the figure:

```{r}
pDHaldane <- 2.87e-9
``` 

The Bayes' factor:

```{r}
BF <- pDHaldane / pDnull
BF
```

### Part D

The BF using the Savage-Dickey method is the ratio of percentage of prior within the ROPE over the percentage of posterior within the ROPE.

```{r}
BFSD <- 0.08 / 1.67
BFSD
```

### Part E


```{r 1_E, cache = T}
# midly informed prior 
a <- 2
b <- 4

BernBeta(c(a, b), c(rep(0, N-z), rep(1, z)) , 
  ROPE = c(0.48, 0.52),
  plotType = "Bars", showCentTend = "Mode", showHDI = TRUE, showpD = TRUE)
```

```{r}
pDInformed <- 2.22e-7

BF <- pDInformed / pDnull
BF
```

### Part F

```{r}
BFSD <- 5 / 1.38 
BFSD
```

### Part G

Haldane prior: HDI for theta is from 0.122 to 0.471 (BF is 3.69).

Informed prior: HDI for theta is from 0.145 to 0.462 (BF is 0.05).


### Part H

The mild-informed prior is preferred for model comparison based on the BF.

## Exercise 12.2

### Model

```{r mod}
OneOddGroupModelComp2E <- function()
{
  out <- list()
  
  oldClass(out) <- "OneOddGroupModelComp2E"
  return(out)
}

mod <- OneOddGroupModelComp2E()
```

### Data

```{r dat}
dataList <- genMCMC(mod, returnDataList = TRUE)

str(dataList)

aggregate(dataList$nCorrOfSubj, by = list(dataList$CondOfSubj) , FUN = mean )
```


### MCMC chain

Generate the MCMC chain:

```{r mcmc, results = 'hide', cache = T}
out <- genMCMC(mod)
```

### Diagnostics

Display diagnostics of chain, for specified parameters:

```{r diag, cache = T}
pnames <- c("mdlIdx","omega[1]","omega0","kappa[1]","theta[1]")
for(p in pnames) { 
  diagMCMC(out, p)
}
```

### Plotting

```{r plot, cache = T}
plotMCMC(mod, out)
```

### Part A

Model 2 (the single-mode model) is preferred with the posterior probability 0.864.

### Part B

The plots of the diferences imply that omega1 is distinct from the others.

### Model with group 1

Generate the MCMC chain:

```{r mcmc2, results = 'hide', cache = T}
out2 <- genMCMC(mod, model = "group1") 
```

```{r plot2, cache = T}
plotMCMC(mod, out2) 
```
