---
title: "Examples from BayesianFirstAid R package based on rjags"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures/", comment = NA, results = 'verbatim', tidy = F, message = F, warning = F, echo = T, cache = TRUE)
```

# About

The idea is to go through tests implemented in [Bayesian First Aid](https://github.com/rasmusab/bayesian_first_aid) R package
and try different Baeysian engines: rjags (used by Bayesian First Aid), rstan and pystan.
So, well-known statistical tests in the Bayesian fashion will help to learn the syntaxes of different tools.
This document shows code for rjags.

As stated in Bayesian First Aid [page](https://github.com/rasmusab/bayesian_first_aid):

Currently Bayesian First Aid implements five alternative models to classical tests. Each of these alternative models are described in detail on my blog (see links below):

* Binomial Test: bayes.binom.test(x, n) . [Code example and details](http://sumsar.net/blog/2014/01/bayesian-first-aid-binomial-test/)
* One Sample and Paired Samples t-test: bayes.t.test(x) . [Code example and details](http://sumsar.net/blog/2014/02/bayesian-first-aid-one-sample-t-test/)
* Two Sample t-test: bayes.t.test(x, y) . Code example and details
* Pearson Correlation Test: bayes.cor.test(x, y) . Code example and details
* Test of Proportions: bayes.prop.test(x, n) . Code example and details
* Poisson test: bayes.poisson.test(x, T) . Code example and details

# Include

```{r, cache = FALSE}
library(BayesianFirstAid)
```

# Binomial Test

```{r binom1}
fitb <- bayes.binom.test(c(9, 2))
fitb
```

```{r plot_fitb}
plot(fitb)
```

Model code in rjags:

```{r code_fitb}
model.code(fitb)
```

## Not a uniform prior

The default flat prior used in the example above.

```{r default_prior_binom}
xp <- seq(0, 1, length = 100)
yp <- dbeta(xp, 1, 1)
plot(xp, yp, t = 'l')
```

Let's try another one `beta(2, 2)`.

```{r prior_binom}
xp <- seq(0, 1, length = 100)
yp <- dbeta(xp, 2, 2)
plot(xp, yp, t = 'l')
```

The `rjags` code is:

```{r binom2}
# Setting up the data
x <- 9 
n <- 11 

# The model string written in the JAGS language
model_string <- "model {
  x ~ dbinom(theta, n)
  theta ~ dbeta(2, 2)
  x_pred ~ dbinom(theta, n)
}"

# Running the model
model <- jags.model(textConnection(model_string), data = list(x = x, n = n), 
                    n.chains = 3, n.adapt=1000)
samples <- coda.samples(model, c("theta", "x_pred"), n.iter=5000)

sf <- BayesianFirstAid:::mcmc_stats(samples)
sf
```

