---
title: "Ex. 9.1"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures-9-1/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T)
```

### Parameters

```{r par}
dir <- "~/git/variani/study/01-book-Kruschke-DBDA/doc/programs/DBDA2Eprograms"
```

## Include DBDA2 files

```{r inc, cache = F}
library(coda)

library(stats) # `aggregate` function
library(graphics) # `par` function
```

### Source files

```{r src, echo = F, results = 'hide'}
dir0 <- getwd()
dir <- "~/git/variani/study/01-book-Kruschke-DBDA/doc/programs/DBDA2Eprograms"
f <- "DBDA2E-utilities.R"

setwd(dir)
source(f)
setwd(dir0)

# for plotting in Rmd by diagMCMC function
openGraph <- function(...) return(invisible())
```

Sourced files: ``r paste(f, collapse = ", ")``.


## Example `Jags-Ydich-XnomSsubj-MbinomBetaOmegaKappa-Example.R`

See code in the `Jags-Ydich-XnomSsubj-MbinomBetaOmegaKappa-Example.R` file.

### Source files

```{r src2, echo = F, results = 'hide'}
dir0 <- getwd()
dir <- "~/git/variani/study/01-book-Kruschke-DBDA/doc/programs/DBDA2Eprograms"
f <- "Jags-Ydich-XnomSsubj-MbinomBetaOmegaKappa.R"

setwd(dir)
source(f)
setwd(dir0)

# for plotting in Rmd by diagMCMC function
openGraph <- function(...) return(invisible())
```

Sourced files: ``r paste(f, collapse = ", ")``.

### Parameters

```{r par1}
sName <- "s"
yName <- "y"
fn <- "Jags-Ydich-XnomSsubj-MbernBetaOmegaKappa-"
 
nsteps <- 20000
```

### Data

Read data:

```{r dat}
dat <- read.csv(file.path(dir, "TherapeuticTouchData.csv"))
str(dat)
```

Create a data list for JAGS:

```{r dat_list}
y <- dat$y
s <- as.numeric(dat$s) # ensures consecutive integer levels

z <- aggregate(y, by = list(s), FUN = sum)$x
N <- aggregate(rep(1,length(y)), by = list(s), FUN = sum)$x
Nsubj <- length(unique(s))

dataList <- list(z = z, N = N, Nsubj = Nsubj)
```  

Print `dataList`:

```{r dataList}
dataList
```
  
### MCMC chain  

Generate the MCMC chain:

```{r mcmc, results = 'hide', cache = TRUE}
myData <- dat
out <- genMCMC(data = myData, sName = sName, yName=yName,
  numSavedSteps = nsteps, saveName = fn, thinSteps = 10)
```

### Plot diagnostics

Display diagnostics of chain, for specified parameters:

```{r diag, cache = T}
pnames <- varnames(out)
for(p in pnames[c(1:2, length(pnames) - 1, length(pnames))]) { 
  diagMCMC(out, p)
}
```  
  
### Plot posteriors
  
```{r post, cache = T}
plotMCMC(out, data = myData, sName = sName, yName = yName, 
  compVal = 0.5 , #rope=c(0.45,0.55) , # Therapeutic touch
  diffIdVec = c(1, 14, 28),              # Therapeutic touch
  compValDiff = 0.0)
```  
  
## Exercise 9.1

```{r kappa_distr_comp}
gammaShRaFromMeanSD(1, 10)
gammaShRaFromModeSD(1, 10)
```

```{r plot_kappa_distr_comp}
xp <- seq(0, 100, by = 0.5)

plot(xp, dgamma(xp, 1.105125, 0.1051249), t = 'l')
lines(xp, dgamma(xp, 0.01, 0.01), col = 'red')

xp <- seq(0.1, 75, by = 0.5)

plot(xp, dgamma(xp, 1.105125, 0.1051249), t = 'l', main = "Between 0.1 and 75: mode-based dominates")
lines(xp, dgamma(xp, 0.01, 0.01), col = 'red')

xp <- seq(0, 0.1, by = 0.5)

plot(xp, dgamma(xp, 1.105125, 0.1051249), t = 'l', main = "Between 0 and 0.1: mean-based dominates")
lines(xp, dgamma(xp, 0.01, 0.01), col = 'red')


xp <- seq(75, 100, by = 0.5)

plot(xp, dgamma(xp, 1.105125, 0.1051249), t = 'l', main = "Between 75 and 100: mean-based dominates")
lines(xp, dgamma(xp, 0.01, 0.01), col = 'red')
```  
  
  
  
  
  
  
  
  
  
  
  
