---
title: "Exercises  from Little Inference Book"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    toc: true
    keep_md: true
---

```{r options, echo = F}
opts_chunk$set(fig.path = "figures/", comment = NA, results = 'markup', tidy = F, message = F, warning = F, echo = T, cache = TRUE)
```

# Include

```{r}
library(pander)
library(tidyverse)
```

## Settings


```{r settings, cache = FALSE}
theme_set(theme_light())

panderOptions('table.style', 'rmarkdown')

panderOptions('table.split.table', Inf)
panderOptions('knitr.auto.asis', FALSE)
```

# Simulations

## The standard error of the mean

What is the variability of the mean of a sample, if we get only a single realization?
[Simulation](https://leanpub.com/LittleInferenceBook/read#leanpub-auto-simulation-example-1-standard-normals)
with repeated sample avergaging can answer the question.

Before going to the code, a small recap:

* The sample variance, $S^2$, estimates the population variance, $\sigma^2$.
* The distribution of the sample variance is centered around $\sigma^2$.
* The variance of the sample mean is $\sigma^2 / n$.

```{r ex1}
nsim <- 1000
n <- 10

set.seed(1)

sim_means <- 
  rep(n, nsim) %>%
  sapply(. %>% rnorm %>% mean)

sim_means_sd <- sim_means %>% sd
```

```{r ex1_tab, echo = F, results = "asis", dependson = "ex1"}
tab <- data.frame(method = c("theory: 1 / sqrt(n)", "simulation: sim_means_sd"), sd = c( 1/sqrt(n), sim_means_sd))
pandoc.table(tab, style = 'rmarkdown')
```

```{r ex1_hist, dependson = "ex1"}
ggplot(data.frame(mean = sim_means), aes(mean)) + geom_histogram() + 
  geom_vline(xintercept = c(-sim_means_sd, sim_means_sd)) 
``` 
 
